---
title: "Gap Minder set"
output:
  html_document: default
date: "2024-06-05"
---

### libraries

```{r}
library(tidyverse)
library(gapminder)
library(shiny)
library(shinydashboard)
library(broom)
library(purrr)
library(DT)
library(flextable)


```

### Data

```{r}

view(gapminder)

```

### Getting a sense of data

```{r}

# Data contains information on 142 countries
unique(gapminder$country)

# Data contains 5 continents: Africa, Americas, Asia, Europe, and Oceania
unique(gapminder$continent)

# Years span from 1972 - 2007
gapminder %>% 
  select(year) %>% 
  arrange(desc(year))

# There are 12 observations per country (where each observation 5 years apart)
gapminder %>% 
  select(country) %>% 
  count(country, sort = TRUE)

# Iceland has the highest average life expectancy while Sierra Leone has the lowest
gapminder %>% 
  group_by(country) %>% 
  summarise(Mean_life = mean(lifeExp)) %>% 
  arrange(desc(Mean_life))

# China has the highest average population size while Sao Tome and Principe has the lowest (Iceland has the second lowest)
gapminder %>% 
  group_by(country) %>% 
  drop_na() %>% 
  summarise(average_population_size = mean(pop),
            IQR = IQR(pop)) %>% 
  arrange(desc(average_population_size))




```

### First graph and table (life expectancy of country by continent)

```{r}

# Visual
gapminder %>% 
  filter(continent == "Asia") %>% 
  group_by(country) %>% 
  summarise(average_life = mean(lifeExp)) %>% 
  mutate(country = fct_reorder(country, average_life)) %>% 
  top_n(25) %>% 
  ggplot() + geom_col(aes(country, average_life, fill = country)) + coord_flip() +
  labs(title = "Life Expectancy Among Countries",
       subtitle = "Asia Continent",
       x = "",
       y = "Average Life Expectancy") +
  theme_bw() +
  theme(legend.position = "none") 

# Table
gapminder %>% 
  filter(continent == "Asia") %>% 
  group_by(country) %>% 
  summarise(average_life = round(mean(lifeExp), 3),
            Median = round(median(lifeExp), 3),
            IQR = round(IQR(lifeExp), 3)) %>% 
  arrange(desc(average_life)) %>% 
  mutate(country = fct_reorder(country, average_life)) %>% 
  top_n(25) 
  
```

# Statistical testing: T.TEST

```{r}

# Filter data from two countries
two_countries <- gapminder %>% 
  filter(country %in% c("Afghanistan", "Albania"))

# Obtain the mean life for each country for geom line visual
average_expectancy <- two_countries %>% 
  group_by(country) %>% 
  summarise(average_life = mean(lifeExp))

# Graph
ggplot(data = two_countries, aes(lifeExp, fill = country)) +
  geom_density(alpha = 0.6) +
  xlim(20,90) +
  theme_bw() +
  geom_vline(data = average_expectancy, aes(xintercept = average_life, color = country), linetype = "dashed")


# subset of data for t.test
country_one <- subset(gapminder, country == "Afghanistan")
country_two <- subset(gapminder, country == "Albania")

# two sample t-test
t_1 <- t.test(country_one$lifeExp, country_two$lifeExp, alternative = "two.sided", conf.level = 0.95)

# Converting results to table
summary <- map_df(list(t_1), tidy)

#prints t.test results
summary


```

# Linear model section (pop vs GDP)

notes:

-   Doesn't appear to be normally distributed
-   Some outliers have

```{r}

# Creating subset
simple_country <- gapminder %>% 
  filter(country == "Afghanistan")

# Creating model
model <- lm(gdpPercap ~ pop, data = simple_country)

# Summary
summary(model)

# Plot linear model
plot(model, which = c(1))

# Creating visual (pop vs gdp)
gapminder %>% 
  filter(country == "Afghanistan") %>% 
  ggplot(aes(x = pop, y = gdpPercap)) + geom_point() + geom_line(color = "red") +
  scale_x_continuous(labels = scales::comma) + 
  scale_y_continuous(labels = scales::comma) +
  theme_bw() +
  labs(title = "Population vs. GDP",
       subtitle = "Country Selected: ",
       x = "Population",
       y = "GDP")

# Second version (year vs gdp)
gapminder %>% 
  filter(country == "Afghanistan") %>% 
  ggplot(aes(x = year, y = gdpPercap)) + geom_point() + geom_line(color = "red") +
  theme_bw() +
  labs(title = "GDP per Year",
       subtitle = "Country Selected: ",
       x = "Year",
       y = "GDP")
  


```

### GDP per country (and by continent)

```{r}

# Visual barplots
gapminder %>% 
  filter(continent == "Asia") %>% 
  group_by(country) %>% 
  summarize(GDP = mean(gdpPercap)) %>% 
  arrange(desc(GDP)) %>% 
  mutate(country = fct_reorder(country, GDP)) %>% 
  top_n(20) %>% 
  ggplot() + geom_col(aes(x = country, y = GDP, fill = country)) + coord_flip() +
  theme_bw() +
  theme(legend.position = "none") +
  labs(title = "Average GDP per Country",
       subtitle = "Chosen Continent: ",
       x = "",
       y = "GDP")

# Table
gapminder %>% 
  filter(continent == "Asia") %>% 
  group_by(country) %>% 
  summarize(GDP = mean(gdpPercap)) %>% 
  top_n(20) %>% 
  arrange(desc(GDP))
 


```

notes:

for linear model:

try to create multiple display visuals of the summary lm plots under each plot on the top. - visually it should look like 4 plots total

### Shiny Dashboard

```{r}

# UI section
ui <- dashboardPage(
  
  dashboardHeader(title = "GapMinder",
                  tags$li(class = "dropdown", tags$a(href = "#", title = "Home", icon("home"), id = "home-link"))),
                  
  dashboardSidebar(
    
    # Home menu tab
    menuItem("Home", tabName = "home", icon = icon("home")),
    
    # First menu tab
    menuItem("Visual Comparisons", tabName = "visuals", icon = icon("chart-simple"),
             menuSubItem("Life Expectancy Comparison", tabName = "life_expectancy_visual"),
             menuSubItem("GDP Comparison", tabName = "GDP_visual" )),
    
    # Second menu tab
    menuItem("Inferential Statistics", tabName = "stats", icon = icon("brain"),
             menuSubItem("Comparision testing", tabName = "testing"),
             menuSubItem("Linear Model", tabName = "linear"))
             
          ),
  
  dashboardBody(

    tabItems(
      
      
#--------------------------- Home page content --------------------------------
    tabItem(tabName = "home",
    fluidRow(
      tabBox(title = "", width = "80%", height = 10,
      tabPanel("Overview of Dashboard",
        collapsible = FALSE,
        HTML("Use this page to navigate efficiently. <ul>
              <li>In the Visual Comparison section, find two tabs:
                <ul>
                  <li><strong>Life Expectancy Comparison:</strong> Compares average life expectancy among 25 countries by continent.</li>
                  <li><strong>GDP Comparison:</strong> Compares average GDP among 25 countries by continent.</li>
                </ul>
              </li>
            </ul>
            
        
            <ul>
              <li>In the Inferential Statistics section, find two tabs:
                <ul>
                  <li><strong>Comparison testing:</strong> This function compares the average life expectancy between any two selected countries and performs a t-test automatically. Note: A p-value below 0.05 indicates a significant difference in average life expectancy between the chosen countries.</li>
                  <li><strong>Linear Model:</strong>This feature generates linear plots illustrating the relationship between GDP and population, as well as the relationship between GDP and year for the selected country. It subsequently presents four plots under each linear plot for additional visualizations.  </li>
                </ul>
              </li>
            </ul>
            Explore our data and visualization tools."
      )),
      
      tabPanel("gapminder data",
               dataTableOutput("data_set")))

    
    )),
      
      
#---------------------- Visual Comparisons Section -----------------------------

      # Life expectancy tab content (Visual Comparisons)
      tabItem(tabName = "life_expectancy_visual",
         fluidRow(
           box(selectInput("continent_choice", "Select Continent", choices = unique(gapminder$continent)))),
         fluidRow(
           box(dataTableOutput("table_1"), width = "100%")),
         fluidRow(
           box(plotOutput("plot_1"), width = "100%"))
         
         ),
      
      
      # GDP tab content (under Visual Comparisons)
      tabItem(tabName = "GDP_visual",
          fluidRow(
           box(selectInput("continent_choice_GDP", "Select Continent", choices = unique(gapminder$continent)))),
          fluidRow(
            box(dataTableOutput("table_GDP"), width = "100%")),
          fluidRow(
            box(plotOutput("plot_GDP"), width = "100%"))
          
        ),

      
      
#--------------------- Comparison Testing Section -----------------------------

      # first tab content (under Comparison testing tab)
      tabItem(tabName = "testing",
              fluidPage(
                fluidRow(
                  box(selectInput("first_country_choice", "Select First Country", choices = unique(gapminder$country))),
                  box(selectInput("second_country_choice", "Select Second Country", choices = unique(gapminder$country)))),
              fluidRow(
                   box(title = "Statistical Test: Two-Sample T.Test", tableOutput("table_2"), width = 12, offset = 3)),
              fluidRow(
                   box(plotOutput("plot_2"),width = "100%")))),
      
      
      
      # Second tab content (under linear analysis tab)
      tabItem(tabName = "linear",
              fluidRow(
                box(selectInput("model_choice_country", "Select Country", choices = unique(gapminder$country)))),
              fluidRow(
                box(plotOutput("plot_3")),
                box(plotOutput("plot_4"))),
              fluidRow(
          
                tabBox(title = " Linear Model plots for Population and GDP",
                  tabPanel("plot 1", plotOutput("linear_plot_summary_1")),
                  tabPanel("plot 2", plotOutput("linear_plot_summary_2")),
                  tabPanel("plot 3", plotOutput("linear_plot_summary_3")),
                  tabPanel("plot 4", plotOutput("linear_plot_summary_4"))),
                 
                tabBox(title = " Linear Model Plots for GDP and Year ",
                  tabPanel("plot 2-1", plotOutput("linear_plot_summary_21")),
                  tabPanel("plot 2-2", plotOutput("linear_plot_summary_22")),
                  tabPanel("plot 2-3", plotOutput("linear_plot_summary_23")),
                  tabPanel("plot 2-4", plotOutput("linear_plot_summary_24")))
                  
                ))
              )))
                   

# Server Section

server <- function(input,output,session){
  
  
  #----------------------- HomePage section -----------------------------------
  
  output$data_set <- renderDataTable({
    
    return(gapminder)
    
    
  })
  
  
  #--------------------- Visualizations section -------------------------------
  
  # Life Expectancy graph
  output$plot_1 <- renderPlot({
    
  gapminder %>% 
    filter(continent == input$continent_choice) %>% 
    group_by(country) %>% 
    summarise(average_life = mean(lifeExp)) %>% 
    mutate(country = fct_reorder(country, average_life)) %>% 
    top_n(20) %>% 
    ggplot() + geom_col(aes(country, average_life, fill = country)) + coord_flip() +
    labs(title = "Life Expectancy Among Countries",
         subtitle = paste(input$continent_choice,"Continent"),
         x = "",
         y = "Average Life Expectancy") +
    theme_bw() +
    theme(legend.position = "none") 
  
  })
  
  # table along with life expectancy graph
  output$table_1 <- renderDataTable({
    
  gapminder %>% 
    filter(continent == input$continent_choice) %>% 
    group_by(country) %>% 
    summarise(average_life = round(mean(lifeExp), 3),
              Median = round(median(lifeExp), 3),
              IQR = round(IQR(lifeExp), 3)) %>% 
    arrange(desc(average_life)) %>% 
    mutate(country = fct_reorder(country, average_life)) %>% 
    top_n(20) 
    
  })
  
  
  # GDP Visuals
  
  output$plot_GDP <- renderPlot({
    
  gapminder %>% 
    filter(continent == input$continent_choice_GDP) %>% 
    group_by(country) %>% 
    summarize(GDP = mean(gdpPercap)) %>% 
    arrange(desc(GDP)) %>% 
    mutate(country = fct_reorder(country, GDP)) %>% 
    top_n(20) %>% 
    ggplot() + geom_col(aes(x = country, y = GDP, fill = country)) + coord_flip() +
    theme_bw() +
    theme(legend.position = "none") +
    labs(title = "Average GDP per Country",
         subtitle = paste("Chosen Continent:", input$continent_choice_GDP),
         x = "",
         y = "GDP")
    
  })
  
  # GDP table
  output$table_GDP <- renderDataTable({
    
  gapminder %>% 
    filter(continent == input$continent_choice_GDP) %>% 
    group_by(country) %>% 
    summarize(GDP = mean(gdpPercap)) %>% 
    top_n(20) %>% 
    arrange(desc(GDP))
      
      
    
  })
  
  
  
#-------------- Comparision Testing and T.test Section ------------------------

  # Second Graph
  output$plot_2 <- renderPlot({
    
# Filter data from two countries
two_countries <- gapminder %>% 
  filter(country %in% c(input$first_country_choice, input$second_country_choice))

# Obtain the mean life for each country for geom line visual
average_expectancy <- two_countries %>% 
  group_by(country) %>% 
  summarise(average_life = mean(lifeExp))

# Graph
ggplot(data = two_countries, aes(lifeExp, fill = country)) +
  geom_density(alpha = 0.4) +
  xlim(20,90) +
  theme_minimal() +
  labs(title = "Life Expectancy Among Two Countries",
       subtitle = paste(input$first_country_choice, "and", input$second_country_choice),
       x = "Life Expectancy",
       y = "") +
  geom_vline(data = average_expectancy, aes(xintercept = average_life, color = country), linetype = "solid", alpha = 1)

  })
  
  # t test results along with second graph
  output$table_2 <- renderTable({
    
  # subset of data for t.test
  country_one <- subset(gapminder, country == input$first_country_choice)
  country_two <- subset(gapminder, country == input$second_country_choice)
  
  if(input$first_country_choice == input$second_country_choice){
    
    print("Can not perform a t.test, please select two different countries")
    
  } else {
    
    t_1 <- t.test(country_one$lifeExp, country_two$lifeExp, alternative = "two.sided", conf.level = 0.95)

    # Converting results to table
    summary <- map_df(list(t_1), tidy)
    
    #prints t.test results
    summary[c("estimate", "estimate1", "estimate2" , "statistic", "p.value")]

  }
  
  })
  
#------------------------ Linear Model Section --------------------------------- 
  
  # Third section for linear model
  output$plot_3 <- renderPlot({
    
  gapminder %>% 
    filter(country == input$model_choice_country) %>% 
    ggplot(aes(x = pop, y = gdpPercap)) + geom_point() + geom_line(color = "red") +
    scale_x_continuous(labels = scales::comma) + 
    scale_y_continuous(labels = scales::comma) +
    theme_bw() +
    labs(title = "Population vs. GDP",
         subtitle = paste("Country Selected:", input$model_choice_country),
         x = "Population",
         y = "GDP")
  })
  
  
  output$plot_4 <- renderPlot({
    
# Second version (year vs gdp)
gapminder %>% 
  filter(country == input$model_choice_country) %>% 
  ggplot(aes(x = year, y = gdpPercap)) + geom_point() + geom_line(color = "red") +
  theme_bw() +
  labs(title = "GDP per Year",
       subtitle = "Country Selected: ",
       x = "Year",
       y = "GDP")
    
  })
            
  # -------------- Plots for linear model summary (section 1) -------------------
  
  # First plot
  output$linear_plot_summary_1 <- renderPlot({
    
    # Creating subset
    simple_country <- gapminder %>% 
        filter(country == input$model_choice_country)

    # Creating model
    model <- lm(gdpPercap ~ pop, data = simple_country)
    
    # plot of model
    plot(model, which = c(1)) + theme_bw()
    
  })
  
  # Second plot
  output$linear_plot_summary_2 <- renderPlot({
    
     # Creating subset
    simple_country <- gapminder %>% 
        filter(country == input$model_choice_country)

    # Creating model
    model <- lm(gdpPercap ~ pop, data = simple_country)
    
    # plot of model
    plot(model, which = c(2))
    
  })
  
  # Third plot
  output$linear_plot_summary_3 <- renderPlot({
    
     # Creating subset
    simple_country <- gapminder %>% 
        filter(country == input$model_choice_country)

    # Creating model
    model <- lm(gdpPercap ~ pop, data = simple_country)
    
    # plot of model
    plot(model, which = c(3))
    
  })
  
  # Fourth plot
  output$linear_plot_summary_4 <- renderPlot({
    
     # Creating subset
    simple_country <- gapminder %>% 
        filter(country == input$model_choice_country)

    # Creating model
    model <- lm(gdpPercap ~ pop, data = simple_country)
    
    # plot of model
    plot(model, which = c(4))
    
  })
  
   # -------------- Plots for linear model summary (section 2) -------------------
  
  # plot 1
  output$linear_plot_summary_21 <- renderPlot({
    
    # Creating subset
    simple_country <- gapminder %>% 
        filter(country == input$model_choice_country)

    # Creating model
    model <- lm(gdpPercap ~ year, data = simple_country)
    
    # plot of model
    plot(model, which = c(1))
    
    
  })
  
  # plot 2
  output$linear_plot_summary_22 <- renderPlot({
    
     # Creating subset
    simple_country <- gapminder %>% 
        filter(country == input$model_choice_country)

    # Creating model
    model <- lm(gdpPercap ~ year, data = simple_country)
    
    # plot of model
    plot(model, which = c(2))
    
    
  })
  
  # plot 3
  output$linear_plot_summary_23 <- renderPlot({
    
     # Creating subset
    simple_country <- gapminder %>% 
        filter(country == input$model_choice_country)

    # Creating model
    model <- lm(gdpPercap ~ year, data = simple_country)
    
    # plot of model
    plot(model, which = c(3))
    
  })
  
  # plot 4
  output$linear_plot_summary_24 <- renderPlot({
    
     # Creating subset
    simple_country <- gapminder %>% 
        filter(country == input$model_choice_country)

    # Creating model
    model <- lm(gdpPercap ~ year, data = simple_country)
    
    # plot of model
    plot(model, which = c(4))
    
    
  })
  
}

# Running Application
shinyApp(ui,server)

```
